---
title: "Reproducible Research: Peer Assessment 1"
author: "Shenay"
date: "June 2, 2019"
output: 
  html_document:
    keep_md: yes
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.path='Figs/', echo = TRUE)
library(ggplot2)
```

## Loading and preprocessing the data

* Load the data
* Process/transform the data if necessary into a format suitable for analysis

```{r}
setwd("C:/Users/sheng/Coursera/RepData_PeerAssessment1")
if (!file.exists("./data")) {dir.create("./data")}
fileURL <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(fileURL, destfile = "./data/activity.zip", mode = "wb")

unzip("./data/activity.zip")
data <- read.csv("activity.csv")

str(data)  
summary(data) 

data$date <- as.Date(data$date, format = "%Y-%m-%d")
```

## What is mean total number of steps taken per day?

* Calculate the total number of steps taken per day
* Make a histogram of the total number of steps taken per day
* Calculate and report the mean and median of total number of steps taken per day

```{r plot1, fig.width=7, fig.height=5}

total_day <- aggregate(x = data$steps, by = list(data$date), FUN = sum, na.rm = TRUE)
colnames(total_day) <- c("date", "steps")
summary(total_day$steps)

hist(total_day$steps, main = "Histogram of Total Number of Steps Taken Per Day",
     xlab = "Steps", border = "royalblue4", col = "royalblue", las = 1)
abline(v = mean(total_day$steps), lwd = 2)

plot(total_day$date, total_day$steps, type = "h", main = "Total Number of Steps Taken Per Day",
     sub = "Note: Period from 2012-10-01 to 2012-11-30", xlab = "", ylab = "Steps", col = rainbow(15), lwd = 2)
abline(h = mean(total_day$steps), col = "gray")
abline(h = median(total_day$steps), col = "gray50")
legend("topleft", legend = c(paste("Mean =", round(mean(total_day$steps))),
                             paste("Median =", round(median(total_day$steps)))))

```


## What is the average daily activity pattern?

* Make a time series plot of the 5-minute interval and the average number of steps taken, averaged across all days
* Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps

```{r plot2, fig.width=7, fig.height=5}

avg_day <- aggregate(x = data$steps, by = list(data$interval), FUN = mean, na.rm = TRUE)
colnames(avg_day) <- c("interval", "avg.steps")
summary(avg_day$avg.steps)

ggplot(avg_day, aes(x = interval)) +
        geom_line(aes(y = avg.steps)) +
        geom_hline(yintercept = mean(avg_day$avg.steps), col = "blue") +
        geom_hline(yintercept = max(avg_day$avg.steps), col = "green") +
        labs(title = "Time Series Plot of Average Number of Steps Taken (Per 5-Min Interval)",
             x = "Intervals", y = "Steps") +
        annotate("label", x = max(avg_day$avg.steps), y = mean(avg_day$avg.steps), label = "Mean") +
        annotate("label", x = max(avg_day$avg.steps), y = max(avg_day$avg.steps), label = "Maximum") +
        theme_classic()

```

## Imputing missing values

* Calculate and report the total number of missing values in the dataset
* Devise a strategy for filling in all of the missing values in the dataset
* Create a new dataset that is equal to the original dataset but with the missing data filled in
* Make a histogram of the total number of steps taken each day. Calculate and report the mean and median total number of steps taken per day

```{r plot3, fig.width=7, fig.height=5}

sum(is.na(data$steps), na.rm = TRUE)
data$steps[which(is.na(data$steps))] <- mean(data$steps, na.rm = TRUE)

na_index <- data.frame(data[which(is.na(data$steps)), ])
na_index$steps <- replace(na_index$steps, na_index$interval %in% avg_day$interval, avg_day$avg.steps)

data_lessna <- data.frame(data[complete.cases(data), ])
new_data <- rbind(na_index, data_lessna)

newtotal_day <- aggregate(x = new_data$steps, by = list(new_data$date), FUN = sum, na.rm = TRUE)
colnames(newtotal_day) <- c("date", "steps")

summary(newtotal_day$steps)

hist(newtotal_day$steps, main = "Histogram of Total Number of Steps Taken Per Day",
     xlab = "Steps", border = "lightskyblue4", col = "lightskyblue", las = 1)

```


## Are there differences in activity patterns between weekdays and weekends?

* Create a new factor variable in the dataset with two levels – “weekday” and “weekend” 
* Make a panel plot containing a time series plot of the 5-minute interval and the average number of steps taken, averaged across all weekday days or weekend days

```{r plot4, fig.width=7, fig.height=5}

wkday <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")

new_data$weekday <- factor((weekdays(new_data$date) %in% wkday),
                         levels = c(TRUE, FALSE),
                         labels = c("weekday", "weekend"))

newavg_day <- aggregate(steps ~ interval + weekday, data = new_data, FUN = mean)
colnames(newavg_day) <- c("interval", "weekday", "avg.steps")

ggplot(newavg_day, aes(x = interval)) +
        geom_line(aes(y = avg.steps)) +
        facet_wrap(~ weekday, ncol = 1) +
        labs(title = "Time Series Plot of Average Number of Steps Taken",
             x = "Interval", y = "Steps") +
        theme_classic()

```

